### 2024-06-04 thesis/input_files/probabilistic.tex
cleaned all problems and warnings

\section{Quantile forest}
Meinshausen \cite{meinshausen2006quantile} extends the idea of random forest \cite{breiman2001random} generalising it, the result is the quantile forest algorithm. Quantile forest allows us to estimate conditional quantiles in a non parametric fashion.
\\
In order to understand this algorithm, it is first necessary to cover the theory of decision trees and random forests.
\subsection{Decision trees}
Decision tree methods partition recursively the feature space in a set of binary rectangles and then fit a simple model in each of those partitions (the most straightforward is fitting just a constant).
To get started, we first split the space into two disjoint regions, then we model the response variable by the mean of the observed predicted variables with associated features falling in that specific region. Our goal is selecting the best features and best split point to achieve the best generalising fit.
For a visualisation consider the Figures \ref{fig:elements_statistical_learning1} and \ref{fig:elements_statistical_learning2}, where the decision tree algorithm is visualised for a regression problem with two independent variables $X_1$ and $X_2$.
\begin{figure}
    \includegraphics[width=0.5\textwidth]{images/elsii1.png}
    \caption{Two-dimensional feature space partitioned by recursive binary splitting \cite{hastie2009elements}}
    \label{fig:elements_statistical_learning1}
  \end{figure}
\\
\begin{figure}
\includegraphics[width=1\textwidth]{images/elsii2.png}
\caption{Partition tree and regression model \cite{hastie2009elements}}
\label{fig:elements_statistical_learning2}
\end{figure}
\\
Suppose now to partition the feature space into M regions $R_1,\dots,R_M$, then the model reads as follows.
\begin{equation}
    f(x)=\sum\limits_{m=1}^{M}c_m\mathbb{I}_{\{x \in R_m\}}
\end{equation}
\\
It follows that the function minimising the sum of squares is the one with $\hat{c}_m=mean(y_i|x_i \in R_m)$.
Finding the best split in terms of minimum sum of squares is computationally infeasible in practice. Therefore, we approximate a solution by approaching the problem in a greedy fashion.
Let $j$ denote the splitting variable and $s$ be the split point we define the two half planes
\begin{equation}
    R_1(j,s)=\{X|X_j\leq s\} \quad R_2(j,s)=\{X|X_j >s\}
\end{equation}
Then we search for the $s$ and $j$ that solve
\begin{equation}
    \min_{j,s}\left[\min_{c_1} \sum\limits_{x_i \in R_1(j,s)}(y_i-c_1)^2+\min_{c_2} \sum\limits_{x_i \in R_2(j,s)}(y_i-c_2)^2\right]
\end{equation}
The inner problem is easy, as already pointed out, we will have 
\begin{equation}
    \hat{c}_1=mean(y_i|x_i \in R_1) \quad \hat{c}_2=mean(y_i|x_i \in R_2)
\end{equation}
For the outer problem, we scan through all the $(j,s)$ tuples and pick the best pair. 
Next, one or both of these regions from the previous step are split into two more regions. We recurse this process until some stopping condition is triggered (max number of branches, max depth of tree, threshold on the mean squared error (MSE), minimum number of observations in each leaf node).
Notice, this being a greedy algorithm implies that our final solution is guaranteed to be just a local optimum not a global one.
Even though their simplicity, these models have proved themselves to be really powerful. See Figure \ref{fig:decision_tree} for an example where decision trees with different hyperparameters are fitted to a sine wave plus noise. The most popular among these models is the CART \cite{breiman2017classification} tree, its name comes from the fact that it can handle both classification and regression problems.
\begin{figure}
    \includegraphics[width=\textwidth]{images/decision_tree.png}
    \caption{Decision tree regression}
    \label{fig:decision_tree}
    % image generated on coolab to be quick
\end{figure}

@@

\subsection{Random forest}
Decision trees are characterised by high variance and low bias, thus, they can benefit extremely from bagging. Furthermore, every decision tree generated through bagging will be identically distributed (i.d.), thus the expectation of an average of $B$ trees is probabilistically equivalent to the expectation of any such tree. As a consequence, the bias will stay fixed since the bias of the bagged estimator is the same as that of each individual tree. 
\\
Consider positively correlated i.d.\ random variables, then the variance of their average is 
\begin{equation}
    \rho \sigma^2+\frac{1-\rho}{B}\sigma^2
\end{equation}
The second term disappears as $B$ increases, while the first term depends heavily on the correlation between bagged trees. Random forest consists in reducing the correlation between the trees by randomly selecting $m$ of the $p$ features as candidates for splitting; $m$ tipically takes a value in the order of $\sqrt{p}$ or even 1 with default value $m=\left\lfloor \frac{p}{3} \right\rfloor$, while a good minimum node size is around five. 
\\
Letting $T(x;\Theta_b)$ be the $b^{th}$ bagged tree where $\Theta_b$ denotes the randomness characterising its splits, cutpoints and terminal node values, we have that the random forest regressor is given by
\begin{equation}
    \hat{f}_{rf}^{B}=\frac{1}{B}\sum\limits_{b=1}^{B}T(x;\Theta_b)
\end{equation}
For a simple visualisation compare Figure \ref{fig:random_forest} with Figure \ref{fig:decision_tree}.
\begin{figure}
    \includegraphics[width=\textwidth]{images/random_forest.png}
    \caption{Random forest regression}
    \label{fig:random_forest}
    % image generated on coolab to be quick
\end{figure}

@@

\begin{definition}
    The dual function associated to the Lagrangian $\mathcal{L}(x,\lambda, \nu)$ is given by $g(\lambda, \nu)=\underset{x}\inf \ \mathcal{L}(x,\lambda, \nu)$
\end{definition}
where $\lambda$ is called the Lagrange multiplier of the optimisation problem. Such dual formulation has an useful property, which is \begin{equation}\label{weak_duality}
    g(\lambda, \nu)\leq p^*
\end{equation}
where $p^*$ is the optimal value of your optimisation problem. In other words $g(\lambda, \nu)$ is a lower bound for the optimal $p^*$.
Consider now a simple Lagrangian $$\mathcal{L}(x,\lambda, \nu)=f(x)+\sum\limits_{i=1}^n \lambda_i r_i(x) +\sum\limits_{i=1}^{n} \nu_i h_i(x)$$
where $r_i(x)$ are inequality constraints while $h_i(x)$ are equality constraints of the problem. Then it can be noted that, the lower bound on $p^*$ is non trivial only when the Lagrange multiplier $\lambda \succeq 0$.
% when lagrange multiplier is negative, we have that g > p
Therefore, the idea is that by maximising the dual function subject to the constraint $\lambda \succeq 0$, we can obtain an approximate or perfect solution to the primal problem.
\\
To explain why we may or may not be able to attain the best solution to the primal problem by maximising its dual, we have to introduce the concept of duality.
\\
We use $d^*$ to denote the optimal value of the Lagrange dual problem; we can think of it as the best lower bound on $p^*$. 
\\
The Inequality \ref{weak_duality} is called weak duality. The difference $p^*-d^*$ is the optimal duality gap. It is the gap between the optimal value of the primal problem and the best lower bound on it that can be obtained from the Lagrange dual function. Moreover, note that the optimal duality gap is always nonnegative.
We say that strong duality holds, when the optimal duality gap is zero. In other words, the Lagrange dual bound is tight.
\\
Constraint qualifications are conditions under which strong duality holds. One of the most popular is Slater's condition.
\begin{proposition}
    Slater's condition reads as
    \begin{equation}\label{slater_condition}
        \begin{aligned}
            \exists x \in \textrm{relint} \ D \ \textrm{s.t.} & \ r_i(x)<0, \quad i=1, \dots, m \\
            & h_i(x)=0
        \end{aligned}
    \end{equation}
    Where relint $D$ is the relative interior of $ D:=\underset{i=0}{\overset{m}{\cap}} \operatorname {dom} (r_{i})$
\end{proposition}
Slater's theorem naturally follows.
\begin{theorem}
    If Slater's condition holds and the problem is convex then strong duality holds.
\end{theorem}
We can now check that our optimisation problem possesses strong duality by checking Slater's condition.
\\
In our case we do not have any equality constraint, so we do not have to worry about the $h_i(x)=0$ term in Equation \ref{slater_condition}. All we have to check is the convexity of our problem and that there exists an $x$ such that $r_i(x)<0$.
For convexity, a sufficient condition is the positive definiteness of $Q$ in the quadratic programming problem 
\begin{equation}
    \begin{aligned}
        \min \quad & x^\intercal Q x+ c^\intercal x \\        
        s.t \quad& Ax\preceq b
    \end{aligned}
\end{equation}
% In our case, equation \ref{eq:kqr_min5}, we have $w^\intercal w$, thus $Q$ is just the identity matrix which satisfies the positive definiteness requirement. 
This condition is easily checked by the fact that kernel matrices are by definition positive semidefinite.
% $\|w\|_{\mathcal{H}}^2= w^\intercal w= \alpha^\intercal K \alpha$, the last element is positive semidefinitive which in turn implies also $w^\intercal w$ is also positive semidefinite.
Therefore, our problem is convex.
Next we check that Slater's condition holds. Considering first the two non negative constraints on $\xi$ and $\xi^*$, we conclude that $\xi$  and $\xi^*$ have to be greater or equal to zero for the existence of an $x$ satisfying Slater's condition. Thus, let us suppose that $0 \leq \xi \leq \mu$ and $0 \leq \xi^* \leq \mu$.
\\

@@

\\
\begin{equation}
    \begin{aligned}
        w=& \Phi^\intercal(\Phi \Phi^\intercal)^{-1} (y-b)
    \end{aligned}
\end{equation}
We then have for any $\xi > 0$ and $\xi^* > 0$  that
\begin{equation}
    \begin{aligned}
        -\xi + y -\Phi\Phi^\intercal(\Phi \Phi^\intercal)^{-1}(y-b)-b<&0
        \\
        -\xi^* - y +\Phi\Phi^\intercal(\Phi \Phi^\intercal)^{-1}(y-b)+b<&0
    \end{aligned}
\end{equation}
Hence, we conclude that our problem satisfies Slater's condition. Therefore, the solution of the dual and primal problems are equivalent.
\\

@@

\\
First, take the dual function from Equation \ref{eq:kqr_min5}.
\begin{equation}
    \begin{aligned}
        g(\lambda, \lambda^*, \nu, \nu^*)= & \inf_{\xi, \xi^*, w, b} \mathcal{L}(w,b,\xi,\xi^*,\lambda, \lambda^*, \nu, \nu^*)
\end{aligned}
\end{equation}
Setting its derivatives to zero
\begin{equation}\label{eq:lagrange_derivatives}
    \begin{cases}
        \frac{\partial \mathcal{L}}{\partial w}=0 \implies w=\Phi^\intercal(\lambda-\lambda^*)
        \\
        \frac{\partial \mathcal{L}}{\partial b}=0 \implies (\lambda-\lambda^*)^\intercal\mathbb{1}=0
        \\
        \frac{\partial \mathcal{L}}{\partial \xi}=0 \implies Cq \mathbb{1}-\lambda- \nu=0
        \\
        \frac{\partial \mathcal{L}}{\partial \xi^*}=0 \implies C(1-q)\mathbb{1} -\lambda^* -\nu^*=0
    \end{cases}
\end{equation}
As pointed out previously, the lower bound resulting from the dual formulation is non trivial only when the Lagrange multipliers $\lambda$ are $\succeq 0$. Looking at the last two equations of the system \ref{eq:lagrange_derivatives}, this implies the following two constraints $\lambda \in [0, Cq\mathbb{1}]$ and $\lambda^*\in [0, C(1-q)\mathbb{1}]$.
\\
Substitute the conditions for an optimum into \ref{eq:kqr_min5}, we obtain the dual formulation.
% nu and nu^* go away because they are multipliers accounting for equality constraint, which in the optimum have to be satisfied hence are zero => h_i \times \nu=0
\begin{equation}
    \begin{split}
        g(\lambda, \lambda^*) = \quad & \hbox{$\xi^\intercal(Cq\mathbb{1} - \lambda - \nu) + (\xi^*)^\intercal(C(1-q)\mathbb{1} - \lambda^* - \nu^*)$} \\
        & \hbox{$- (\lambda - \lambda^*)^\intercal \Phi \Phi^\intercal (\lambda - \lambda^*)$} \\
        & \hbox{$+ (\lambda - \lambda^*)^\intercal y - (\lambda - \lambda^*)^\intercal b + \frac{1}{2} (\lambda - \lambda^*)^\intercal \Phi \Phi^\intercal (\lambda - \lambda^*)$} \\
        \\
        g(\lambda, \lambda^*) = \quad & \hbox{$0 + 0 - \frac{1}{2} (\lambda - \lambda^*)^\intercal \Phi \Phi^\intercal (\lambda - \lambda^*)$} \\
        & \hbox{$+ (\lambda - \lambda^*)^\intercal y - 0$} \\
        \\
        g(\lambda, \lambda^*) = \quad & \hbox{$- \frac{1}{2} (\lambda - \lambda^*)^\intercal \Phi \Phi^\intercal (\lambda - \lambda^*)$} \\
        & \hbox{$+ (\lambda - \lambda^*)^\intercal y$}
    \end{split}
\end{equation}
Defining $\alpha=(\lambda-\lambda^*)$ and letting $K$ be the kernel matrix, we have that the dual optimisation problem reads as follows
\begin{equation}\label{eq:kqr_min6}
    \begin{aligned}
        \max_{\alpha} \quad & -\frac{1}{2}\alpha^\intercal K\alpha+\alpha^\intercal y\\
    \textrm{s.t.} \quad & 
    C(q-1)\mathbb{1}\preceq \alpha \preceq Cq\mathbb{1}\\
    &\alpha^\intercal\mathbb{1}=0
    \end{aligned}
    \end{equation}
Switching sign, we rephrase it as a minimisation problem, which is the common practice in convex optimisation.
\begin{equation}\label{eq:kqr_min7}
    \begin{aligned}
        \min_{\alpha} \quad & +\frac{1}{2}\alpha^\intercal K \alpha-\alpha^\intercal y\\
    \textrm{s.t.} \quad & 
    C(q-1)\mathbb{1}\preceq \alpha \preceq Cq\mathbb{1}\\
    &\alpha^\intercal\mathbb{1}=0
    \end{aligned}
    \end{equation}
The kernel quantile regression estimator is then given by
\begin{equation}
    f(x)=\sum\limits_{i=1}^{n} \alpha_i k(x_i, x)+b
\end{equation}
Since our optimisation problem possesses strong duality and it is differentiable in both the objective and the constraint, we have that it must satisfy the Karush Kuhn Tucker conditions (KKT), see \cite[Section 5.5.3]{boyd2004convex}.
Thanks to the KKT conditions on the primal optimisation problem we have that $f(x_i)=y_i \ \mathrm{for} \ \alpha_i \not \in \{C(q-1), Cq\}$. 
To see this, we have to consider the KKT conditions.
% optimal lambda times optimal x
\begin{equation}
    \begin{aligned}
    \lambda r_i(x)=&0, \quad i=1,\dots,n,
    \\
    \nabla \mathcal{L}(x)=&0
\end{aligned}
\end{equation}
In our setting we have
\begin{equation}
    \begin{aligned}
        \lambda_i(\xi_i-y_i+r_i)=&0
        \\
        \lambda_i^*(\xi_i^*+y_i-r_i)=&0
        \\
        \nu_i \xi_i=&0
        \\
        \nu_i^* \xi_i^*=&0
        \\
        \nabla \mathcal{L} =&0
    \end{aligned}
\end{equation}
Using the gradient of the Lagrangian of Equation \ref{eq:lagrange_derivatives}, we end up with
\begin{equation}
    \begin{aligned}
        \lambda_i(\xi_i-y_i+r_i)=&0
        \\
        \lambda_i^*(\xi_i^*+y_i-r_i)=&0
        \\
        (Cq-\lambda_i) \xi_i=&0
        \\
        (C(1-q)-\lambda_i) \xi_i^*=&0
    \end{aligned}
\end{equation}

>>>>

\section{Quantile forest}
Meinshausen \cite{meinshausen2006quantile} extends the idea of random forest \cite{breiman2001random} generalising it, the result is the quantile forest algorithm. Quantile forest allows us to estimate conditional quantiles in a non parametric fashion.
In order to understand this algorithm, it is first necessary to cover the theory of decision trees and random forests.
\subsection{Decision trees}
Decision tree methods partition recursively the feature space in a set of binary rectangles and then fit a simple model in each of those partitions (the most straightforward is fitting just a constant).
To get started, we first split the space into two disjoint regions, then we model the response variable by the mean of the observed predicted variables with associated features falling in that specific region. Our goal is selecting the best features and best split point to achieve the best generalising fit.
For a visualisation consider the Figures \ref{fig:elements_statistical_learning1} and \ref{fig:elements_statistical_learning2}, where the decision tree algorithm is visualised for a regression problem with two independent variables $X_1$ and $X_2$.
\begin{figure}
    \includegraphics[width=0.5\textwidth]{images/elsii1.png}
    \caption{Two-dimensional feature space partitioned by recursive binary splitting \cite{hastie2009elements}}
    \label{fig:elements_statistical_learning1}
  \end{figure}

\begin{figure}
\includegraphics[width=1\textwidth]{images/elsii2.png}
\caption{Partition tree and regression model \cite{hastie2009elements}}
\label{fig:elements_statistical_learning2}
\end{figure}

Suppose now to partition the feature space into M regions $R_1,\dots,R_M$, then the model reads as follows.
\begin{equation}
    f(x)=\sum\limits_{m=1}^{M}c_m\mathbb{I}_{\{x \in R_m\}}
\end{equation}

It follows that the function minimising the sum of squares is the one with $\hat{c}_m=mean(y_i|x_i \in R_m)$.
Finding the best split in terms of minimum sum of squares is computationally infeasible in practice. Therefore, we approximate a solution by approaching the problem in a greedy fashion.
Let $j$ denote the splitting variable and $s$ be the split point we define the two half planes
\begin{equation}
    R_1(j,s)=\{X|X_j\leq s\} \quad R_2(j,s)=\{X|X_j >s\}
\end{equation}
Then we search for the $s$ and $j$ that solve
\begin{equation}
    \min_{j,s}\left[\min_{c_1} \sum\limits_{x_i \in R_1(j,s)}(y_i-c_1)^2+\min_{c_2} \sum\limits_{x_i \in R_2(j,s)}(y_i-c_2)^2\right]
\end{equation}
The inner problem is easy, as already pointed out, we will have 
\begin{equation}
    \hat{c}_1=mean(y_i|x_i \in R_1) \quad \hat{c}_2=mean(y_i|x_i \in R_2)
\end{equation}
For the outer problem, we scan through all the $(j,s)$ tuples and pick the best pair. 
Next, one or both of these regions from the previous step are split into two more regions. We recurse this process until some stopping condition is triggered (max number of branches, max depth of tree, threshold on the mean squared error (MSE), minimum number of observations in each leaf node).
Notice, this being a greedy algorithm implies that our final solution is guaranteed to be just a local optimum not a global one.
Even though their simplicity, these models have proved themselves to be really powerful. See Figure \ref{fig:decision_tree} for an example where decision trees with different hyperparameters are fitted to a sine wave plus noise. The most popular among these models is the CART \cite{breiman2017classification} tree, its name comes from the fact that it can handle both classification and regression problems.
\begin{figure}
    \includegraphics[width=\textwidth]{images/decision_tree.png}
    \caption{Decision tree regression}
    \label{fig:decision_tree}
    % image generated on coolab to be quick
\end{figure}

@@

\subsection{Random forest}
Decision trees are characterised by high variance and low bias, thus, they can benefit extremely from bagging. Furthermore, every decision tree generated through bagging will be identically distributed (i.d.), thus the expectation of an average of $B$ trees is probabilistically equivalent to the expectation of any such tree. As a consequence, the bias will stay fixed since the bias of the bagged estimator is the same as that of each individual tree. 
\\
Consider positively correlated i.d.\ random variables, then the variance of their average is 
\begin{equation}
    \rho \sigma^2+\frac{1-\rho}{B}\sigma^2
\end{equation}
The second term disappears as $B$ increases, while the first term depends heavily on the correlation between bagged trees. Random forest consists in reducing the correlation between the trees by randomly selecting $m$ of the $p$ features as candidates for splitting; $m$ tipically takes a value in the order of $\sqrt{p}$ or even 1 with default value $m=\left\lfloor \frac{p}{3} \right\rfloor$, while a good minimum node size is around five.

Letting $T(x;\Theta_b)$ be the $b^{th}$ bagged tree where $\Theta_b$ denotes the randomness characterising its splits, cutpoints and terminal node values, we have that the random forest regressor is given by
\begin{equation}
    \hat{f}_{rf}^{B}=\frac{1}{B}\sum\limits_{b=1}^{B}T(x;\Theta_b)
\end{equation}
For a simple visualisation compare Figure \ref{fig:random_forest} with Figure \ref{fig:decision_tree}.
\begin{figure}
    \includegraphics[width=\textwidth]{images/random_forest.png}
    \caption{Random forest regression}
    \label{fig:random_forest}
    % image generated on coolab to be quick
\end{figure}

@@

\begin{definition}
    The dual function associated to the Lagrangian $\mathcal{L}(x,\lambda, \nu)$ is given by $g(\lambda, \nu)=\underset{x}\inf \ \mathcal{L}(x,\lambda, \nu)$
\end{definition}
where $\lambda$ is called the Lagrange multiplier of the optimisation problem. Such dual formulation has an useful property, which is \begin{equation}\label{weak_duality}
    g(\lambda, \nu)\leq p^*
\end{equation}
where $p^*$ is the optimal value of your optimisation problem. In other words $g(\lambda, \nu)$ is a lower bound for the optimal $p^*$.
Consider now a simple Lagrangian $$\mathcal{L}(x,\lambda, \nu)=f(x)+\sum\limits_{i=1}^n \lambda_i r_i(x) +\sum\limits_{i=1}^{n} \nu_i h_i(x)$$
where $r_i(x)$ are inequality constraints while $h_i(x)$ are equality constraints of the problem. Then it can be noted that, the lower bound on $p^*$ is non trivial only when the Lagrange multiplier $\lambda \succeq 0$.
% when lagrange multiplier is negative, we have that g > p
Therefore, the idea is that by maximising the dual function subject to the constraint $\lambda \succeq 0$, we can obtain an approximate or perfect solution to the primal problem.
To explain why we may or may not be able to attain the best solution to the primal problem by maximising its dual, we have to introduce the concept of duality.
We use $d^*$ to denote the optimal value of the Lagrange dual problem; we can think of it as the best lower bound on $p^*$. 
The Inequality \ref{weak_duality} is called weak duality. The difference $p^*-d^*$ is the optimal duality gap. It is the gap between the optimal value of the primal problem and the best lower bound on it that can be obtained from the Lagrange dual function. Moreover, note that the optimal duality gap is always nonnegative.
We say that strong duality holds, when the optimal duality gap is zero. In other words, the Lagrange dual bound is tight.

Constraint qualifications are conditions under which strong duality holds. One of the most popular is Slater's condition.
\begin{proposition}
    Slater's condition reads as
    \begin{equation}\label{slater_condition}
        \begin{aligned}
            \exists x \in \textrm{relint} \ D \ \textrm{s.t.} & \ r_i(x)<0, \quad i=1, \dots, m \\
            & h_i(x)=0
        \end{aligned}
    \end{equation}
    Where relint $D$ is the relative interior of $ D:=\underset{i=0}{\overset{m}{\cap}} \operatorname {dom} (r_{i})$
\end{proposition}
Slater's theorem naturally follows.
\begin{theorem}
    If Slater's condition holds and the problem is convex then strong duality holds.
\end{theorem}
We can now check that our optimisation problem possesses strong duality by checking Slater's condition.

In our case we do not have any equality constraint, so we do not have to worry about the $h_i(x)=0$ term in Equation \ref{slater_condition}. All we have to check is the convexity of our problem and that there exists an $x$ such that $r_i(x)<0$.
For convexity, a sufficient condition is the positive definiteness of $Q$ in the quadratic programming problem 
\begin{equation}
    \begin{aligned}
        \min \quad & x^\intercal Q x+ c^\intercal x \\        
        s.t \quad& Ax\preceq b
    \end{aligned}
\end{equation}
% In our case, equation \ref{eq:kqr_min5}, we have $w^\intercal w$, thus $Q$ is just the identity matrix which satisfies the positive definiteness requirement. 
This condition is easily checked by the fact that kernel matrices are by definition positive semidefinite.
% $\|w\|_{\mathcal{H}}^2= w^\intercal w= \alpha^\intercal K \alpha$, the last element is positive semidefinitive which in turn implies also $w^\intercal w$ is also positive semidefinite.
Therefore, our problem is convex.
Next we check that Slater's condition holds. Considering first the two non negative constraints on $\xi$ and $\xi^*$, we conclude that $\xi$  and $\xi^*$ have to be greater or equal to zero for the existence of an $x$ satisfying Slater's condition. Thus, let us suppose that $0 \leq \xi \leq \mu$ and $0 \leq \xi^* \leq \mu$.

@@

\begin{equation}
    \begin{aligned}
        w=& \Phi^\intercal(\Phi \Phi^\intercal)^{-1} (y-b)
    \end{aligned}
\end{equation}
We then have for any $\xi > 0$ and $\xi^* > 0$  that
\begin{equation}
    \begin{aligned}
        -\xi + y -\Phi\Phi^\intercal(\Phi \Phi^\intercal)^{-1}(y-b)-b<&0
        \\
        -\xi^* - y +\Phi\Phi^\intercal(\Phi \Phi^\intercal)^{-1}(y-b)+b<&0
    \end{aligned}
\end{equation}
Hence, we conclude that our problem satisfies Slater's condition. Therefore, the solution of the dual and primal problems are equivalent.

@@

First, take the dual function from Equation \ref{eq:kqr_min5}.
\begin{equation}
    \begin{aligned}
        g(\lambda, \lambda^*, \nu, \nu^*)= & \inf_{\xi, \xi^*, w, b} \mathcal{L}(w,b,\xi,\xi^*,\lambda, \lambda^*, \nu, \nu^*)
\end{aligned}
\end{equation}
Setting its derivatives to zero
\begin{equation}\label{eq:lagrange_derivatives}
    \begin{cases}
        \frac{\partial \mathcal{L}}{\partial w}=0 \implies w=\Phi^\intercal(\lambda-\lambda^*)
        \\
        \frac{\partial \mathcal{L}}{\partial b}=0 \implies (\lambda-\lambda^*)^\intercal\mathbb{1}=0
        \\
        \frac{\partial \mathcal{L}}{\partial \xi}=0 \implies Cq \mathbb{1}-\lambda- \nu=0
        \\
        \frac{\partial \mathcal{L}}{\partial \xi^*}=0 \implies C(1-q)\mathbb{1} -\lambda^* -\nu^*=0
    \end{cases}
\end{equation}
As pointed out previously, the lower bound resulting from the dual formulation is non trivial only when the Lagrange multipliers $\lambda$ are $\succeq 0$. Looking at the last two equations of the system \ref{eq:lagrange_derivatives}, this implies the following two constraints $\lambda \in [0, Cq\mathbb{1}]$ and $\lambda^*\in [0, C(1-q)\mathbb{1}]$.

Substitute the conditions for an optimum into \ref{eq:kqr_min5}, we obtain the dual formulation.
% nu and nu^* go away because they are multipliers accounting for equality constraint, which in the optimum have to be satisfied hence are zero => h_i \times \nu=0
\begin{equation}
    \begin{split}
        g(\lambda, \lambda^*) = \quad & \hbox{$\xi^\intercal(Cq\mathbb{1} - \lambda - \nu) + (\xi^*)^\intercal(C(1-q)\mathbb{1} - \lambda^* - \nu^*)$} \\
        & \hbox{$- (\lambda - \lambda^*)^\intercal \Phi \Phi^\intercal (\lambda - \lambda^*)$} \\
        & \hbox{$+ (\lambda - \lambda^*)^\intercal y - (\lambda - \lambda^*)^\intercal b + \frac{1}{2} (\lambda - \lambda^*)^\intercal \Phi \Phi^\intercal (\lambda - \lambda^*)$} \\
        \\
        g(\lambda, \lambda^*) = \quad & \hbox{$0 + 0 - \frac{1}{2} (\lambda - \lambda^*)^\intercal \Phi \Phi^\intercal (\lambda - \lambda^*)$} \\
        & \hbox{$+ (\lambda - \lambda^*)^\intercal y - 0$} \\
        \\
        g(\lambda, \lambda^*) = \quad & \hbox{$- \frac{1}{2} (\lambda - \lambda^*)^\intercal \Phi \Phi^\intercal (\lambda - \lambda^*)$} \\
        & \hbox{$+ (\lambda - \lambda^*)^\intercal y$}
    \end{split}
\end{equation}
Defining $\alpha=(\lambda-\lambda^*)$ and letting $K$ be the kernel matrix, we have that the dual optimisation problem reads as follows
\begin{equation}\label{eq:kqr_min6}
    \begin{aligned}
        \max_{\alpha} \quad & -\frac{1}{2}\alpha^\intercal K\alpha+\alpha^\intercal y\\
    \textrm{s.t.} \quad & 
    C(q-1)\mathbb{1}\preceq \alpha \preceq Cq\mathbb{1}\\
    &\alpha^\intercal\mathbb{1}=0
    \end{aligned}
    \end{equation}
Switching sign, we rephrase it as a minimisation problem, which is the common practice in convex optimisation.
\begin{equation}\label{eq:kqr_min7}
    \begin{aligned}
        \min_{\alpha} \quad & +\frac{1}{2}\alpha^\intercal K \alpha-\alpha^\intercal y\\
    \textrm{s.t.} \quad & 
    C(q-1)\mathbb{1}\preceq \alpha \preceq Cq\mathbb{1}\\
    &\alpha^\intercal\mathbb{1}=0
    \end{aligned}
    \end{equation}
The kernel quantile regression estimator is then given by
\begin{equation}
    f(x)=\sum\limits_{i=1}^{n} \alpha_i k(x_i, x)+b
\end{equation}
Since our optimisation problem possesses strong duality and it is differentiable in both the objective and the constraint, we have that it must satisfy the Karush Kuhn Tucker conditions (KKT), see \cite[Section 5.5.3]{boyd2004convex}.
Thanks to the KKT conditions on the primal optimisation problem we have that $f(x_i)=y_i \ \mathrm{for} \ \alpha_i \not \in \{C(q-1), Cq\}$. 
To see this, we have to consider the KKT conditions.
% optimal lambda times optimal x
\begin{equation}
    \begin{aligned}
    \lambda r_i(x)=&0, \quad i=1,\dots,n,
    \\
    \nabla \mathcal{L}(x)=&0
\end{aligned}
\end{equation}
In our setting we have
\begin{equation}
    \begin{aligned}
        \lambda_i(\xi_i-y_i+r_i)=&0
        \\
        \lambda_i^*(\xi_i^*+y_i-r_i)=&0
        \\
        \nu_i \xi_i=&0
        \\
        \nu_i^* \xi_i^*=&0
        \\
        \nabla \mathcal{L} =&0
    \end{aligned}
\end{equation}
Using the gradient of the Lagrangian of Equation \ref{eq:lagrange_derivatives}, we end up with
\begin{equation}
    \begin{aligned}
        \lambda_i(\xi_i-y_i+r_i)=&0
        \\
        \lambda_i^*(\xi_i^*+y_i-r_i)=&0
        \\
        (Cq-\lambda_i) \xi_i=&0
        \\
        (C(1-q)-\lambda_i) \xi_i^*=&0
    \end{aligned}
\end{equation}
